---
title: "Analysis Document for Shizuka et al., Global analysis in a widespread songbird reveals that phenotypic integration and modularity evolve along with plumage coloration"
output: html_document
date: "2025-01-09"
---

# Load Libraries
```{r}
require(pacman)
p_load(tidyverse,qgraph,igraph,devtools,patchwork,ggrepel,glue,gtools,PHENIX,dplyr,rsample,pbapply,parallel,lme4,broom, cowplot)

```

# Functions

## Bootstrap procedure

BOOTSTRAP PROCEDURE

Does the following:
  
  1. Runs analysis over the traits specified by 'columns', with IDs in 'id'
  
  2. Iterates this whole procedure over iterate_over column (default="population")
  
  3. Bootstraps data 'boot_n' times (default=1e4), balancing sample sizes over 'strata_var' (default="sex")
      *'strata_levels'= case-insensitive 2 levels of strata,
                        ordered Female first (default=c("f","m"))
    ** Bootstraps are parallelized, change cores with the 'cor_num' parameter.
       default: cor_num=parallel::detectCores()-1
  
  4. Calculates averages for traits specified in 'columns' within each bootstrap sample
  
  5. Calculates the corrected phenotypic integration (PINT.c) value using PHENIX::pint()
  
  6. Calculates dichromatism (or dimorphism) for traits specified in 'columns' and sex defined by 'strata' using dimorphCal()
  
  7. Calculates correlations of 'cor_type' (default = "p"; passed to cor()) between:
      a. (PI) and trait mean for "toi"
      b. (PI) and dichromatism across trait(s) in 'toi'
  
  NOTES:
  *   'keep_cols' refers to columns you want to keep in your outputs
  **  'seed' is set to 99 by default, allowing for repeatable bootstrap results.

```{r}
# Function Definitions -----------------------------------------------------
boot_analy <- function(df = NULL,
                       columns = NULL,
                       toi = "r.chrom",
                       boot_n = 1000,
                       strata_var = "sex",
                       strata_levels = c("f", "m"),
                       id_col = "band",
                       keep_cols=NULL,
                       iterate_over = "population",
                       cor_type = "p",
                       seed = 99,
                       cor_num=parallel::detectCores()-1) {
  
  set.seed(seed) #to make this repeatable
  if (is.null(df)) {
    message("must input your data as the 'df' parameter")
  }
  if (is.null(columns)) {
    message("specify what columns you want to analyze (i.e. traits to measure dichromatism for)")
  }
  #make case-insensitive levels for filtering
  stratum1 <- c(toupper(strata_levels[1]), tolower(strata_levels[1]))
  stratum2 <- c(toupper(strata_levels[2]), tolower(strata_levels[2]))
 
  #And test just as an extra precaution
  sum_levels_match_expected<-sum(unlist(unique(df[,substitute(strata_var)])) %in% strata_levels)
  
  if(sum_levels_match_expected!=2){
    stop("Levels don't match. Make sure strata_levels=c(val1,val2), found in your data frame.")
  }
  
  #Get a vector of unique entries to iterate over
  uniques <-
    unique(df[, iterate_over]) %>% na.omit() %>% unlist() %>% as.vector()
  
  
  # Iterate analysis over all populations/subgroups -------------------------
  all_subgroups <- pbapply::pblapply(1:length(uniques), function(i) {
    subgroup <- uniques[i]
    message("\n\n# Analyzing ",
            iterate_over,
            " ",
            i,
            " of ",
            length(uniques),
            ": '",
            subgroup,
            "'")
    # Get the subgroup (i.e. subset with population==x)
    d_i <-
      df %>% dplyr::filter(eval(as.symbol(iterate_over)) == subgroup)
    
    #Now do bootstrapping with each subgroup_i, stratified by 'strata'
    boots <- rsample::bootstraps(d_i, boot_n, strata = strata_var)
    
    #######
    # ANALYZE each bootstrap, within population i -----------------------------
    message("   Bootstrap Progress")
    boot_results <- pbapply::pblapply(1:length(boots$splits),cl = parallel::detectCores()-1,FUN= function(ii) {
      
      boot_ii <- boots$splits[[ii]] %>%
        rsample::analysis() %>%
        select(all_of(c(id_col,strata_var,iterate_over,columns,keep_cols))) %>%  #we want the value of these vars
        group_by(!!sym(strata_var))#man this escaping w/ !! vs {{ }} distinction is annoying!
      
      # A) Trait Means
      trait_means_ii <- boot_ii  %>% 
        dplyr::summarise(across(where(is.double), mean, na.rm = TRUE)) %>%
        dplyr::rename_with(~ paste0("avg_", .), where(is.double))
      #preserve keep_cols
      if(!is.null(keep_cols)){
        trait_means_ii <- bind_cols(boot_ii[1:2,keep_cols],trait_means_ii)
      }
      
      # B) Sex Differences in Traits
      
      sex_diffs_ii <-
        dimorphCal(boot_ii,
                   columns,
                   strata_var = strata_var,
                   strata_levels = strata_levels,
                   keep_cols=keep_cols)%>% 
        rename_with(~paste0("SD_",.),all_of(columns))
      
      
      # C) Phenotypic Integration across all traits
      PI_ii <- lapply(strata_levels, function(stratum) {
        
        boot_ii_stratum <- boot_ii %>%
          dplyr::filter(eval(as.symbol(strata_var)) == stratum) %>% ungroup %>%  select(all_of(columns))
        
        PI_ii <- PHENIX::pint(na.omit(boot_ii_stratum))
        dplyr::tibble({
          {
            strata_var
          }
        } := stratum,
        N_PINT = PI_ii$N,
        PINT = PI_ii$PINT,
        PINT.c = PI_ii$PINT.c)
      }) %>% dplyr::bind_rows()
      
      
      # D) Aggregate summary of PI and trait avgs
      
      summary_ii <-PI_ii %>% 
        mutate({{iterate_over}} := subgroup, boot_i = ii) %>%
        relocate({{iterate_over}}, boot_i) %>%
        left_join(x = ., y = trait_means_ii, by = {{strata_var}})
      # E) Return summary data
      list(
        summary = summary_ii,
        sex_diffs = as_tibble(sex_diffs_ii) %>% mutate({
          {
            iterate_over
          }
        } := subgroup, boot_i = ii) %>%
          relocate({
            {
              iterate_over
            }
          }, boot_i)
      )
      
    })  # End bootstraps within population_i
    
    #Combine sex diffs and summary table across all
    boot_summ <-
      lapply(boot_results, function(x)
        x$summary) %>% bind_rows()
    boot_sex_diffs <-
      lapply(boot_results, function(x)
        x$sex_diffs) %>% bind_rows()
    
    
    subgroup_output <- list(boot_summ = boot_summ ,
                            boot_sex_diffs = boot_sex_diffs)
    
    subgroup_output
    
  })# End lapply over all populations/subgroups
  
  #Organize final output
  boot_summ_tibble <-
    lapply(1:length(all_subgroups), function(i) {
      all_subgroups[[i]]$boot_summ
    }) %>% bind_rows
  boot_sex_diffs_tibble <-
    lapply(1:length(all_subgroups), function(i) {
      all_subgroups[[i]]$boot_sex_diffs
    }) %>% bind_rows
  
  mean_traits_tibble <- boot_summ_tibble %>% select(-boot_i) %>%
    group_by(!!sym(iterate_over), !!sym(strata_var),!!!syms(keep_cols)) %>% 
    summarize_if(is.double, mean, na.rm =TRUE)
  
  mean_sex_diffs_tibble <-
    boot_sex_diffs_tibble %>% select(-boot_i) %>%
    group_by(!!sym(iterate_over),!!!syms(keep_cols)) %>% 
    summarize(across(all_of(paste0("SD_",columns)), mean, na.rm=TRUE))
  
  #calculate correlations bw phenotypic integration and Trait Means for Trait(s) of Interest
  cor_trait_avgs_df <- lapply(1:length(toi), function(i) {
    lapply(1:boot_n, function(j) {
      #stratum1 is usually going to be female (the first entry in 'strata_levels')
      df_j1 <-
        boot_summ_tibble %>% dplyr::filter(boot_i == j &
                                             eval(as.symbol(strata_var)) %in% stratum1)
      df_j2 <-
        boot_summ_tibble %>% dplyr::filter(boot_i == j &
                                             eval(as.symbol(strata_var)) %in% stratum2)
      #usually the female correlation
      #Note, we're using the PINT.c value, which corrects for number of individuals and traits 
      #for each population
      cor_1 <-
        cor(df_j1$PINT.c ,
            df_j1[, paste0("avg_", toi[i])] %>% unlist(),
            method = cor_type,
            use = "complete.obs")
      #usually the male correlation
      cor_2 <-
        cor(df_j2$PINT.c ,
            df_j2[, paste0("avg_", toi[i])] %>% unlist(),
            method = cor_type,
            use = "complete.obs")
      tibble({
        {
          strata_var
        }
      } := strata_levels, trait_x1 = "PINT.c", trait_x2 = paste0("avg_", toi[i]), cor =
        c(cor_1, cor_2))
    }) %>% bind_rows
  }) %>% bind_rows
  
  #Collapse correlations across bootstrap correlations for trait means
  #Just temporarily disabling this b/c of a persistent message
  options(dplyr.summarise.inform = F)
  cor_trait_avgs_tibble <- cor_trait_avgs_df %>%
    group_by(!!sym(strata_var), trait_x1, trait_x2) %>% summarize(
      mean_cor = mean(cor, na.rm = T),
      ci_low = quantile(
        cor,
        probs = .025,
        na.rm = T,
        names = F,
        type = 7
      ),
      ci_high = quantile(
        cor,
        probs = .975,
        na.rm = T,
        names = F,
        type = 7
      )
    ) %>% arrange(trait_x2)
  
  
  #Calculate correlations between phenotypic integration and sex difference in toi(s)
  cor_sex_diff_df <- lapply(1:length(toi), function(i) {
    lapply(1:boot_n, function(j) {
      #stratum1 is usually going to be female (the first entry in 'strata_levels')
      sex_diff_j <-
        boot_sex_diffs_tibble %>% dplyr::filter(boot_i == j)
      df_j1 <-
        boot_summ_tibble %>% dplyr::filter(boot_i == j &
                                             eval(as.symbol(strata_var)) %in% stratum1)
      df_j2 <-
        boot_summ_tibble %>% dplyr::filter(boot_i == j &
                                             eval(as.symbol(strata_var)) %in% stratum2)
      #usually the female correlation
      cor_1 <- cor(sex_diff_j[, paste0("SD_",toi[i])] %>% unlist() ,
                   df_j1$PINT.c,
                   method = cor_type,
                   use = "complete.obs")
      #usually the male correlation
      cor_2 <- cor(sex_diff_j[,paste0("SD_",toi[i])] %>% unlist() ,
                   df_j2$PINT.c,
                   method = cor_type,
                   use = "complete.obs")
      tibble({
        {
          strata_var
        }
      } := strata_levels, trait_x1 = "PINT.c", trait_x2 = paste0("sex_diff_", toi[i]), cor =
        c(cor_1, cor_2))
    }) %>% bind_rows
  }) %>% bind_rows
  
  #Collapse correlations across bootstrap correlations for trait means
  cor_trait_sex_diff_tibble <- cor_sex_diff_df %>%
    group_by(!!sym(strata_var), trait_x1, trait_x2) %>%
    summarize(
      mean_cor = mean(cor, na.rm = T),
      ci_low = quantile(
        cor,
        probs = .025,
        na.rm = T,
        names = F,
        type = 7
      ),
      ci_high = quantile(
        cor,
        probs = .975,
        na.rm = T,
        names = F,
        type = 7
      )
    ) %>%
    arrange(trait_x2)
  options(dplyr.summarise.inform = T)#turn messaging back on
  
  message(":) ALL DONE!\n")
  SUMMARY <- list(
    boot_n= boot_n,
    boot_summ = boot_summ_tibble,
    boot_sex_diffs = boot_sex_diffs_tibble,
    mean_traits = mean_traits_tibble,
    mean_sex_diffs = mean_sex_diffs_tibble,
    cor_pint_x_avg_trait = cor_trait_avgs_tibble,
    cor_pint_x_sex_diff = cor_trait_sex_diff_tibble
  )
  
  
}
```


```{r, eval=F}

#Import all data
d00<-read_csv("Data/all_populations.csv")
nrow(d00)

traits<-c('tail.mean','t.avg.bright','t.hue','t.chrom','r.avg.bright','r.hue','r.chrom','b.avg.bright','b.hue','b.chrom','v.avg.bright','v.hue','v.chrom')
#Just the color traits
traits_col <- traits[-c(1)]

#limit analysis to pops with at least N samples
(pop_summary<-d00 %>% group_by(population,sex) %>%  summarise(n=n()) %>%  pivot_wider(names_from=sex,values_from=n,names_prefix = "n_") %>% mutate(n_TOT=n_F+n_M) %>% as.data.frame())

#Let's say 12 is our minimum number of each sex
min_samples<-12
pops_w_min_samples<-pop_summary %>% filter(n_F>=min_samples & n_M>=min_samples)
pops_w_20_samples<-pop_summary %>% filter(n_F>=20 & n_M>=20)
d0<-d00 %>% filter(population %in% pops_w_min_samples$population)
nrow(d0)
d0$population<-as.factor(d0$population)
d0$sex<-as.factor(d0$sex)

#Let's only work with Colorado samples from 2008 (before many experiments)
d<-d0 %>% filter(population!="colorado"|population=="colorado"&year==2008)
#Now CO has a more comparable N to other pops
d %>% group_by(population,sex) %>%  summarise(n=n()) %>%  pivot_wider(names_from=sex,values_from=n,names_prefix = "n_") %>% mutate(n_TOT=n_F+n_M) %>% as.data.frame()

#Export info on just the populations we're using &rename Taiwan to Taipei, Taiwan for consistencey
d %>% select(population,location,year,lat,long,hybrid_zone,zone) %>%
  mutate(location=case_when(location=="Taiwan"~"Taipei, Taiwan",.default=location)) %>%
  distinct(population,.keep_all = T) %>% write_csv(.,file="Data/populations_analyzed.csv")

write.csv(d, "Data/data_for_submission.csv")
```

# Data filtering

Here is how we get from all data to the final dataset we use, which includes 28 populations (not run here).
```{r, eval=F}

#Import all data
d00<-read_csv("Data/all_populations.csv")
nrow(d00)

traits<-c('tail.mean','t.avg.bright','t.hue','t.chrom','r.avg.bright','r.hue','r.chrom','b.avg.bright','b.hue','b.chrom','v.avg.bright','v.hue','v.chrom')
#Just the color traits
traits_col <- traits[-c(1)]

#limit analysis to pops with at least N samples
(pop_summary<-d00 %>% group_by(population,sex) %>%  summarise(n=n()) %>%  pivot_wider(names_from=sex,values_from=n,names_prefix = "n_") %>% mutate(n_TOT=n_F+n_M) %>% as.data.frame())

#Let's say 12 is our minimum number of each sex
min_samples<-12
pops_w_min_samples<-pop_summary %>% filter(n_F>=min_samples & n_M>=min_samples)
pops_w_20_samples<-pop_summary %>% filter(n_F>=20 & n_M>=20)
nrow(pops_w_min_samples) #28 populations with at least 12 individuals
d0<-d00 %>% filter(population %in% pops_w_min_samples$population)
nrow(d0)
d0$population<-as.factor(d0$population)
d0$sex<-as.factor(d0$sex)

#Let's only work with Colorado samples from 2008 (before many experiments)
d<-d0 %>% filter(population!="colorado"|population=="colorado"&year==2008)
#Now CO has a more comparable N to other pops
d %>% group_by(population,sex) %>%  summarise(n=n()) %>%  pivot_wider(names_from=sex,values_from=n,names_prefix = "n_") %>% mutate(n_TOT=n_F+n_M) %>% as.data.frame()

#Export info on just the populations we're using &rename Taiwan to Taipei, Taiwan for consistencey
d %>% select(population,location,year,lat,long,hybrid_zone,zone) %>%
  mutate(location=case_when(location=="Taiwan"~"Taipei, Taiwan",.default=location)) %>%
  distinct(population,.keep_all = T) %>% write_csv(.,file="Data/populations_analyzed.csv")

write.csv(d, "Data/data_for_submission.csv")

```

# Running the bootstrap procedure

Here is how we ran the bootstrapping procedure using the function above (not run here).
```{r, eval=F}
d=read.csv("Data/data_for_submission.csv")

# VERY Time consuming!
# Uncomment if you want to run the full 10e4 bootstraps

res <- boot_analy(
  df = d,
  columns = traits_col,
  toi = c("r.avg.bright","r.chrom","b.chrom","b.avg.bright", "t.chrom"),
  boot_n = 1e4,
  strata_var="sex",
  strata_levels = c("F", "M"),
  keep_cols = c("location", "lat", "long")
  )
```

# Start here

## Import data

### The bootstrap results
```{r}
res<-readRDS("/Users/dshizuka2/Dropbox/Dai_Research/Main Projects/BARS_phenotypicintegration/results_10k_bootstraps.RDS")
```

### Raw data (not bootstrapped)
```{r}
d=read.csv("../Data/data_for_submission.csv")
```

### Fst data
```{r}
fst<-read.csv("../Data/pairwise_Fst/pairwise_Fst_table.csv")
```

Create a dataset that only includes populations with Fst data
```{r}
pops_w_fst<-fst$population
d_gen<-res$boot_sum %>% filter(population %in% pops_w_fst)
```

### Phenotypic Integration analysis

These are the analyses that Table 1 is built from.

Sset up the dataset that includes the Fst data.
```{r}
d_gen<-left_join(d_gen,fst[,c("population","weighted_Fst")],by="population") %>% select(population,location,sex,boot_i,PINT,PINT.c,avg_r.chrom,avg_t.chrom,avg_b.chrom,avg_v.chrom,weighted_Fst)
d_gen_pops<-d_gen %>% distinct(population) %>% unlist
d_gen_m<-d_gen %>% filter(sex=="M")
d_gen_f<-d_gen %>% filter(sex=="F")
```

```{r, eval=F}
#Run all of the models and save the results so that we don't have to run them each time.

lm_results_r.chrom_fst <- pbapply::pblapply(1:max(d_gen$boot_i), function(i) {
  pops_boot_i <- d_gen %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_r.chrom+ weighted_Fst+as.factor(sex),data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_r.chrom=mod$estimate[2],est_Fst=mod$estimate[3],est_sexM=mod$estimate[4])
}) %>% bind_rows

lm_results_t.chrom_fst <- pbapply::pblapply(1:max(d_gen$boot_i), function(i) {
  pops_boot_i <- d_gen %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_t.chrom+ weighted_Fst+as.factor(sex),data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_t.chrom=mod$estimate[2],est_Fst=mod$estimate[3],est_sexM=mod$estimate[4])
}) %>% bind_rows

#Running model with sex as a covariate and belly chroma
lm_results_b.chrom_fst <- pbapply::pblapply(1:max(d_gen$boot_i), function(i) {
  pops_boot_i <- d_gen %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_b.chrom+ weighted_Fst+as.factor(sex),data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_b.chrom=mod$estimate[2],est_Fst=mod$estimate[3],est_sexM=mod$estimate[4])
}) %>% bind_rows

#Running model with sex as a covariate and vent chroma
lm_results_v.chrom_fst <- pbapply::pblapply(1:max(d_gen$boot_i), function(i) {
  pops_boot_i <- d_gen %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_v.chrom+ weighted_Fst+as.factor(sex),data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_v.chrom=mod$estimate[2],est_Fst=mod$estimate[3],est_sexM=mod$estimate[4])
}) %>% bind_rows

save(lm_results_r.chrom_fst, lm_results_t.chrom_fst, lm_results_b.chrom_fst, lm_results_v.chrom_fst, file="Data/PINTanalysis_results_w_Fst.rds")
```

```{r}
load("../Data/PINTanalysis_results_w_Fst.rds")
```

Results for breast color, as reported in paper
```{r}
# This is the result reported in paper
#95CIs for analysis with sex as covariate, rather than splitting up analyses
#Significant effect of breast darkness on phenotypic integration (PCIT)
quantile(lm_results_r.chrom_fst$est_avg_r.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_r.chrom_fst$est_avg_r.chrom,na.rm=T,names=F)
#Significant effect of population genetics (Fst) on phenotypic integration (PCIT)
quantile(lm_results_r.chrom_fst$est_Fst,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_r.chrom_fst$est_Fst,na.rm=T,names=F)
#Nonsignificant effect of sex on phenotypic integration (PCIT)
quantile(lm_results_r.chrom_fst$est_sexM,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_r.chrom_fst$est_sexM,na.rm=T)

#lm_results_r.chrom%>% pivot_longer(cols=starts_with("est_")) %>% ggplot()+geom_histogram(aes(x=value),fill="gray60",col="gray30")+facet_wrap(~name,ncol = 1)+geom_vline(xintercept=0,col="royalblue")+labs(title="Effect estimates",subtitle="PINT.c~ R_chrom + sex (for 10^4) bootstraps")


#Running FST model with sex as a covariate and with throat



# This is the result reported in paper
#95CIs for analysis with sex as covariate, rather than splitting up analyses
#Significant effect of breast darkness on phenotypic integration (PCIT)
quantile(lm_results_t.chrom_fst$est_avg_t.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_t.chrom_fst$est_avg_t.chrom,na.rm=T,names=F)
#Significant effect of population genetics (Fst) on phenotypic integration (PCIT)
quantile(lm_results_t.chrom_fst$est_Fst,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_t.chrom_fst$est_Fst,na.rm=T,names=F)
#Nonsignificant effect of sex on phenotypic integration (PCIT)
quantile(lm_results_t.chrom_fst$est_sexM,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_t.chrom_fst$est_sexM,na.rm=T)




# This is the result reported in paper
#95CIs for analysis with sex as covariate, rather than splitting up analyses
#Significant effect of breast darkness on phenotypic integration (PCIT)
quantile(lm_results_b.chrom_fst$est_avg_b.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_b.chrom_fst$est_avg_b.chrom,na.rm=T,names=F)
#Significant effect of population genetics (Fst) on phenotypic integration (PCIT)
quantile(lm_results_b.chrom_fst$est_Fst,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_b.chrom_fst$est_Fst,na.rm=T,names=F)
#Nonsignificant effect of sex on phenotypic integration (PCIT)
quantile(lm_results_b.chrom_fst$est_sexM,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_b.chrom_fst$est_sexM,na.rm=T)



# This is the result reported in paper
#95CIs for analysis with sex as covariate, rather than splitting up analyses
#Significant effect of breast darkness on phenotypic integration (PCIT)
quantile(lm_results_v.chrom_fst$est_avg_v.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_v.chrom_fst$est_avg_v.chrom,na.rm=T,names=F)
#Significant effect of population genetics (Fst) on phenotypic integration (PCIT)
quantile(lm_results_v.chrom_fst$est_Fst,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_v.chrom_fst$est_Fst,na.rm=T,names=F)
#Nonsignificant effect of sex on phenotypic integration (PCIT)
quantile(lm_results_v.chrom_fst$est_sexM,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_results_v.chrom_fst$est_sexM,na.rm=T)
```

###supplemental analysis: do this for all 28 populations by excluding genomic data


Build the dataset:
```{r}

boot_f<-res$boot_sum %>% filter(sex=="F") %>% select(population, location, boot_i, sex,avg_r.chrom,avg_t.chrom,avg_b.chrom,avg_v.chrom,PINT.c)
boot_m<-res$boot_sum %>% filter(sex=="M") %>% select(population, location, boot_i, sex,avg_r.chrom,avg_t.chrom,avg_b.chrom,avg_v.chrom,PINT.c)
boot_both<-rbind(boot_f,boot_m)

```

Run all analyses and save results (not run):
```{r, eval=F}

boot_f<-res$boot_sum %>% filter(sex=="F") %>% select(population, location, boot_i, sex,avg_r.chrom,avg_t.chrom,avg_b.chrom,avg_v.chrom,PINT.c)
boot_m<-res$boot_sum %>% filter(sex=="M") %>% select(population, location, boot_i, sex,avg_r.chrom,avg_t.chrom,avg_b.chrom,avg_v.chrom,PINT.c)
boot_lm<-rbind(boot_f,boot_m)

#Run models with sex as a factor
lm_res_R<- pbapply::pblapply(1:max(boot_both$boot_i), function(i) {
  pops_boot_i <- boot_both %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_r.chrom+sex,data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_r.chrom=mod$estimate[2], est_sex=mod$estimate[3])
}) %>% bind_rows

#Throat Chroma
lm_res_T<- pbapply::pblapply(1:max(boot_both$boot_i), function(i) {
  pops_boot_i <- boot_both %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_t.chrom+sex,data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_t.chrom=mod$estimate[2], est_sex=mod$estimate[3])
}) %>% bind_rows

#Belly Chroma
lm_res_B<- pbapply::pblapply(1:max(boot_both$boot_i), function(i) {
  pops_boot_i <- boot_both %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_b.chrom+sex,data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_b.chrom=mod$estimate[2], est_sex=mod$estimate[3])
}) %>% bind_rows

#Vent Chroma
lm_res_V<- pbapply::pblapply(1:max(boot_both$boot_i), function(i) {
  pops_boot_i <- boot_both %>% filter(boot_i == i)
  mod<-lm(PINT.c~avg_v.chrom+sex,data=pops_boot_i) %>% tidy()
  out<-tibble(boot=i,est_avg_v.chrom=mod$estimate[2], est_sex=mod$estimate[3])
}) %>% bind_rows


save(lm_res_R, lm_res_T, lm_res_B, lm_res_V, file="Data/PINTanalysis_results_no_Fst.rds")

```

```{r}
load("../Data/PINTanalysis_results_no_Fst.rds")
```
Breast color
```{r}

#Significant effect of chroma
quantile(lm_res_R$est_avg_r.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_R$est_avg_r.chrom,na.rm=T)

#Nonsignificant effect of sex
quantile(lm_res_R$est_sex,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_R$est_sex,na.rm=T)

```

Throat color:
```{r}
#Run both in a model with sex as a factor


#Significant effect of chroma
quantile(lm_res_T$est_avg_t.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_T$est_avg_t.chrom,na.rm=T)

#Nonsignificant effect of sex
quantile(lm_res_T$est_sex,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_T$est_sex,na.rm=T)
```

Belly color
```{r}
# Look at effect of belly on PINT -----------------------------
#Run both in a model with sex as a factor

#Significant effect of chroma
quantile(lm_res_B$est_avg_b.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_B$est_avg_b.chrom,na.rm=T)

#Nonsignificant effect of sex
quantile(lm_res_B$est_sex,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_B$est_sex,na.rm=T)
```

Vent color
```{r}
# Look at effect of vent on PINT -----------------------------
#Run both in a model with sex as a factor

#Significant effect of chroma
quantile(lm_res_V$est_avg_v.chrom,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_V$est_avg_v.chrom,na.rm=T)

#Nonsignificant effect of sex
quantile(lm_res_V$est_sex,probs=c(.025,.975),na.rm=T,names=F,type=7)
mean(lm_res_V$est_sex,na.rm=T)
```


####Figure 3
```{r}

#Make males show up on the left to be consistent with other figures.
G_r_simple<-res$mean_traits %>%  mutate(sex=factor(sex, levels=c("M", "F"))) %>%
    ggplot(aes(x = avg_r.chrom, y = PINT.c)) +
    geom_point(size=3,pch=21,col="black", aes(fill = avg_r.chrom)) +
    scale_fill_gradient(
      limits = range(res$mean_traits$avg_r.chrom),
      low = "#FFFFCC",
      high = "#CC6600",
      guide = "none"
    ) + 
    geom_smooth(method="lm", color="black") +
    facet_wrap( ~ sex) + 
    xlab("Breast | Average Population Color (Chroma)")+
    ylab("Phenotypic Integration") +
    theme_bw() +
  ylim(0.75, 3.0) +
    theme(strip.text=element_blank(),axis.text=element_text(size=14), axis.title=element_text(size=16), panel.grid.minor=element_blank(), panel.grid.major=element_blank()) 

G_t_simple<-res$mean_traits %>%  mutate(sex=factor(sex, levels=c("M", "F"))) %>%
  ggplot(aes(x = avg_t.chrom, y = PINT.c)) +
  geom_point(size=3,pch=21,col="black", aes(fill = avg_t.chrom)) +
  scale_fill_gradient(
    limits = range(res$mean_traits$avg_t.chrom),
    low = "#FFFFCC",
    high = "#CC6600",
    guide = "none"
  ) + 
  geom_smooth(method="lm", color="black") +
  facet_wrap( ~ sex) + 
  xlab("Throat | Average Population Color (Chroma)")+
  ylab("Phenotypic Integration") +
  theme_bw() +
  ylim(0.75, 3.0) +
  theme(strip.text=element_blank(),axis.text=element_text(size=14), axis.title=element_text(size=16), panel.grid.minor=element_blank(), panel.grid.major=element_blank()) 

plot_grid(G_r_simple, G_t_simple, nrow=2)

```


### Modularity
```{r}
d$population<-as.factor(d$population)
d$sex<-as.factor(d$sex)
### Figure 4A: run community detection on each matrix to quantitatively determine good "modules"
#Male data subset by population
data_list_males<-lapply(levels(d$population),function(x) (subset(d,population==x&sex=="M")))
names(data_list_males)<-levels(d$population)

#Female data subset by population
data_list_females<-lapply(levels(d$population),function(x) (subset(d,population==x&sex=="F")))
names(data_list_females)<-levels(d$population)
```

```{r}
traits<-c('tail.mean','t.avg.bright','t.hue','t.chrom','r.avg.bright','r.hue','r.chrom','b.avg.bright','b.hue','b.chrom','v.avg.bright','v.hue','v.chrom')
traits_col <- traits[-c(1)]

#Male correlations by population
corr_list_males<-lapply(names(data_list_males), function(x) cor(as.matrix(data_list_males[[x]][,traits_col]),method="s",use="pairwise.complete"))
names(corr_list_males)<-levels(d$population)

#Female correlations by population
corr_list_females<-lapply(names(data_list_females), function(x) cor(as.matrix(data_list_females[[x]][,traits_col]),method="s",use="pairwise.complete"))
names(corr_list_females)<-levels(d$population)
```

```{r}
nets_male=lapply(corr_list_males, function(x) {
  absmat=abs(x)
  diag(absmat)=0
  graph_from_adjacency_matrix(absmat, "undirected", weighted=T)
})

nets_female=lapply(corr_list_females, function(x) {
  absmat=abs(x)
  diag(absmat)=0
  graph_from_adjacency_matrix(absmat, "undirected", weighted=T)
})
```

```{r}
## just shorthand for now, removing lower 20% of correlations. Need to figure out a package to use for filtering now that PCIT is defunct.
clusters_male=lapply(nets_male, function(x) {
  g=delete.edges(x, which(E(x)$weight<quantile(E(x)$weight, probs=0.2)))
  cluster_fast_greedy(g, weights=E(g)$weight)
})

clusters_female=lapply(nets_female, function(x) {
  g=delete.edges(x, which(E(x)$weight<quantile(E(x)$weight, probs=0.2)))
  cluster_fast_greedy(g, weights=E(g)$weight)
})

memberships_male=lapply(clusters_male, membership)
comembers_male=lapply(memberships_male, function(x) outer(x, x, "==")+0)

memberships_female=lapply(clusters_female, membership)
comembers_female=lapply(memberships_female, function(x) outer(x, x, "==")+0)
```

```{r}
library(abind)
comembers_male_array=abind(comembers_male, along=3)
comembers_female_array=abind(comembers_female, along=3)

sum_mat_male=apply(comembers_male_array, c(1,2), sum)
sum_mat_male

sum_mat_female=apply(comembers_female_array, c(1,2), sum)
```

```{r}
map.data_male=data.frame(expand.grid(rownames(sum_mat_male), colnames(sum_mat_male)), expand.grid(sum_mat_male))
names(map.data_male)=c("Rows", "Columns", "Values")

matrixplot_male=ggplot(map.data_male, aes(x=Rows, y=Columns, fill=Values)) + 
  geom_tile() +
  scale_fill_gradient(low="white", high="red") +
  theme(legend.position="none", axis.text.x=element_blank(), axis.title=element_blank()) 

map.data_female=data.frame(expand.grid(rownames(sum_mat_female), colnames(sum_mat_female)), expand.grid(sum_mat_female))
names(map.data_female)=c("Rows", "Columns", "Values")

matrixplot_female=ggplot(map.data_female, aes(x=Rows, y=Columns, fill=Values)) + 
  geom_tile() +
  scale_fill_gradient(low="white", high="red") +
  theme(legend.position="none", axis.text.x=element_blank(), axis.title=element_blank())

plot_grid(matrixplot_male, matrixplot_female, nrow=1)


```

### Avg correlations
```{r}
#throat vs. others
patches=c(rep(1, 3), rep(2, 9))
same.patch=outer(patches, patches, FUN="==")
same.patch
patch.names=c("Throat", "Breast-Belly-Vent")
modules=matrix(nrow=length(patches), ncol=length(patches))
for(i in 1:2){
  modules[which(patches==i), which(patches==i)] = i
}
modules
#now take the list of correlation matrices for males across populations and calculate average correlation coefficients within modules (1-4) and between modules (originating from module 1-4)
# mods.list.male=lapply(corr_list_males, function(x) {
#   diag(x)=NA
#   wi_mod1=mean(abs(x[which(modules==1)]), na.rm=T)
#   wi_mod2=mean(abs(x[which(modules==2)]), na.rm=T)
#   wi_mod3=mean(abs(x[which(modules==3)]), na.rm=T)
#   wi_mod4=mean(abs(x[which(modules==4)]), na.rm=T)
#   btw_mod=mean(abs(x[which(is.na(modules))]))
#   btw_12=mean(abs(x[1:3, 4:6]))
#   data.frame(wi_mod1, wi_mod2, wi_mod3, wi_mod4, btw_mod, btw_12)
# })

#throat vs. not throat
mods.list.male=lapply(corr_list_males, function(x) {
  diag(x)=NA
  wi_mod1=mean(abs(x[which(modules==1)]), na.rm=T)
  wi_mod2=mean(abs(x[which(modules==2)]), na.rm=T)
  wi_mod_both=mean(abs(x[which(is.na(modules)==FALSE)]), na.rm=T)
  btw_mod=mean(abs(x[which(is.na(modules))]))
  data.frame(wi_mod1, wi_mod2, wi_mod_both, btw_mod)
})
#organize results into dataframe
mods.dat.male=tibble(bind_rows(mods.list.male))
mods.dat.male$sex="M"
mods.dat.male$population=names(corr_list_males)

#now do the same for female
# mods.list.female=lapply(corr_list_females, function(x) {
#   diag(x)=NA
#   wi_mod1=mean(abs(x[which(modules==1)]), na.rm=T)
#   wi_mod2=mean(abs(x[which(modules==2)]), na.rm=T)
#   wi_mod3=mean(abs(x[which(modules==3)]), na.rm=T)
#   wi_mod4=mean(abs(x[which(modules==4)]), na.rm=T)
#   btw_mod=mean(abs(x[which(is.na(modules))]))
#   btw_12=mean(abs(x[1:3, 4:6]))
#   data.frame(wi_mod1, wi_mod2, wi_mod3, wi_mod4, btw_mod, btw_12)
# })

#throat vs. not throat
mods.list.female=lapply(corr_list_females, function(x) {
  diag(x)=NA
  wi_mod1=mean(abs(x[which(modules==1)]), na.rm=T)
  wi_mod2=mean(abs(x[which(modules==2)]), na.rm=T)
  wi_mod_both=mean(abs(x[which(modules==1|modules==2)]), na.rm=T)
  btw_mod=mean(abs(x[which(is.na(modules))]))
  data.frame(wi_mod1, wi_mod2, wi_mod_both, btw_mod)
})

mods.dat.female=tibble(bind_rows(mods.list.female))
mods.dat.female$sex="F"
mods.dat.female$population=names(corr_list_females)


#Make data frame for main figure 
integ0<-d %>% group_by(population, sex) %>% summarise_at(c("t.chrom","r.chrom","t.avg.bright","r.avg.bright", "b.chrom", "v.chrom", "lat"),mean,na.rm=TRUE) %>% 
  arrange(sex,population) %>% 
  rename(mean.t.chrom=t.chrom,mean.r.chrom=r.chrom,mean.t.avg.bright=t.avg.bright,mean.r.chrom=r.chrom,mean.r.avg.bright=r.avg.bright, mean.b.chrom=b.chrom, mean.v.chrom=v.chrom, latitude=lat)


#now combine the population-level color data with modularity data
mods.dat=bind_rows(mods.dat.female, mods.dat.male)
integ=mods.dat %>% left_join(., integ0) 


```

```{r}
dat2=integ %>% select(wi_mod1, wi_mod2, btw_mod, ends_with("chrom"), sex, population) %>% 
  rename(wi_t=wi_mod1, wi_r=wi_mod2, btw=btw_mod) %>%
  pivot_longer(-c(starts_with("mean"),sex, population), names_to="edge.type", values_to="edge.weight") %>%
  mutate(wi_btw=str_sub(edge.type, start=1, end=2)) %>%
  mutate(wi_btw = replace(wi_btw, wi_btw=="wi", 1)) %>%
  mutate(wi_btw = replace(wi_btw, wi_btw=="bt", 2)) %>%
  mutate(patch=str_sub(edge.type, start=4, end=6)) %>%
  mutate(patch = replace(patch, patch=="t", "within throat")) %>%
  mutate(patch = replace(patch, patch=="r", "within other patches")) %>%
  mutate(patch = replace(patch, patch=="", "between modules")) 


### Build Figure 4C,D
modplot1m=ggplot(dat2 %>% filter(sex=="M"), aes(x=mean.t.chrom, y=edge.weight, fill=patch, color=patch)) +
  geom_smooth( method="lm", se=F, mapping=aes(linetype=wi_btw)) +
  geom_point(alpha=0.8, pch=21, color="black")+
  scale_color_viridis_d(direction=-1)+
  scale_fill_viridis_d(direction=-1)+
  xlim(0.469, 0.581) +
  ylim(0,0.7) +
  theme_cowplot() +
  ylab("Average edge weight") +
  xlab("Average throat chroma of population") +
  ggtitle("Male") +
  guides(linetype=FALSE)

modplot1m_nolegend=modplot1m + theme(legend.position="none")

modplot1f=ggplot(dat2 %>% filter(sex=="F"), aes(x=mean.t.chrom, y=edge.weight, color=patch, fill=patch)) +
  geom_smooth( method="lm", se=F, mapping=aes(linetype=wi_btw)) +
  geom_point(alpha=0.8, pch=21, color="black")+
  scale_color_viridis_d(direction=-1)+
  scale_fill_viridis_d(direction=-1)+
  xlim(0.469, 0.581) +
  ylim(0,0.7) +
  theme_cowplot() +
  ylab("Average edge weight") +
  xlab("Average throat chroma of population") +
  theme(legend.position="none") +
  ggtitle("Female") +
  guides(linetype=FALSE)


modplot2m=ggplot(dat2 %>% filter(sex=="M"), aes(x=mean.r.chrom, y=edge.weight, color=patch, fill=patch)) +
  geom_smooth( method="lm", se=F, mapping=aes(linetype=wi_btw)) +
  geom_point(alpha=0.8, pch=21, color="black")+
  scale_color_viridis_d(direction=-1)+
  scale_fill_viridis_d(direction=-1)+
  theme_cowplot() +
  ylim(0,0.7) +
  ylab("Average edge weight") +
  xlab("Average breast chroma of population") +
  theme(legend.position="none") +
  ggtitle("Male") +
  guides(linetype=FALSE) 

modplot2f=ggplot(dat2 %>% filter(sex=="F"), aes(x=mean.r.chrom, y=edge.weight, color=patch, fill=patch)) +
  geom_smooth( method="lm", se=F, mapping=aes(linetype=wi_btw)) +
  geom_point(alpha=0.8, pch=21, color="black")+
  scale_color_viridis_d(direction=-1)+
  scale_fill_viridis_d(direction=-1)+
  theme_cowplot() +
  ylim(0,0.7) +
  ylab("Average edge weight") +
  xlab("Average breast chroma of population") +
  theme(legend.position="none") +
  ggtitle("Female") +
  guides(linetype=FALSE) 

legend_plot=get_legend(modplot1m)

plot_grid(modplot1m_nolegend, modplot2m, NULL, modplot1f, modplot2f, legend_plot, nrow=2, rel_widths=c(2,2,1))

```
